---
title: "Assignment 12 r prog"
author: "Holden Blank"
date: "2025-04-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
```

The Blankman package provides tools for performing the Friedman test, a non-parametric test for analyzing differences across multiple related samples. This document outlines the primary functions being developed, including their purpose, usage, inputs, outputs, and examples.

Function 1:
Performs the Friedman test on a tidy data frame containing repeated measures or block design data.

```{r}
run_friedman_test <- function(data, subject_col, treatment_col, response_col) {
  formula <- as.formula(paste(response_col, "~", treatment_col, "|", subject_col))
  result <- stats::friedman.test(formula, data = data)

  summary_table <- data %>%
    group_by(.data[[treatment_col]]) %>%
    summarise(Mean_Rank = mean(rank(-.data[[response_col]])), .groups = "drop")

  list(
    statistic = result$statistic,
    p_value = result$p.value,
    df = result$parameter,
    method = result$method,
    summary = summary_table
  )
}
```

Function 2:
Visualizes the average ranks of treatments from the Friedman test using a bar plot.

```{r}
plot_friedman_ranks <- function(data, subject_col, treatment_col, response_col) {
  ranked_data <- data %>%
    group_by(.data[[subject_col]]) %>%
    mutate(Rank = rank(-.data[[response_col]])) %>%
    ungroup()

  summary_ranks <- ranked_data %>%
    group_by(.data[[treatment_col]]) %>%
    summarise(Mean_Rank = mean(Rank), .groups = "drop")

  ggplot(summary_ranks, aes(x = .data[[treatment_col]], y = Mean_Rank)) +
    geom_col() +
    labs(title = "Mean Ranks by Treatment", x = "Treatment", y = "Mean Rank") +
    theme_minimal()
}
```

Function 3:
Generates a tidy summary table of mean ranks and standard deviations by treatment group.
```{r}
get_friedman_summary <- function(data, subject_col, treatment_col, response_col) {
  ranked_data <- data %>%
    group_by(.data[[subject_col]]) %>%
    mutate(Rank = rank(-.data[[response_col]])) %>%
    ungroup()

  ranked_data %>%
    group_by(.data[[treatment_col]]) %>%
    summarise(
      Mean_Rank = mean(Rank),
      SD_Rank = sd(Rank),
      .groups = "drop"
    )
}
```

Notes
All functions require tidy (long-format) data. A helper function could reshape wide -> long.

The plotting function uses descending ranks to reflect higher scores.

Future versions may include post-hoc comparison support (e.g., Nemenyi test).

Testing edge cases with small sample sizes is important to ensure stability.

Package will be extended with vignettes and unit tests.

